---
title: "Exploring Global Fertility: A Journey Across Borders"
subtitle: "Deciphering the Numbers: A Global Perspective on Birthrates and Beyond"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(wbstats)
library(patchwork)

theme_set(theme_minimal())

# Get data from World Bank
str(wb_cachelist, max.level = 1)
fert_tbl <- wb_data("SP.DYN.TFRT.IN", start_date = 1900, end_date = 2020, country = "all")
# Convert year to date object
fert_tbl <- fert_tbl |> 
  mutate(date = ymd(parse_date_time(date, "y")))
```


## Let's Explore the World Ferility Rates

In a world bustling with over **seven billion people**, each individual represents a unique thread in the intricate tapestry of humanity. But beyond the surface of bustling streets and crowded cities lies a story told not through words, but through numbers - the **story of fertility rates**. This vital statistic, often overlooked, serves as a silent narrator of our times, speaking volumes about **societal trends, economic conditions, and cultural shifts.**

As we embark on a journey across continents, we delve into the fascinating world of fertility rates. From the snow-capped villages of Scandinavia to the sun-drenched lands of Sub-Saharan Africa, we'll explore how these numbers shape nations, influence policies, and reflect the diverse tapestry of human existence.

Join me as we unravel the tales hidden within these figures, **discovering not just the "how" and "what," but the profound "why" behind the fertility rates** of various countries. We'll look at how these rates are more than just numbers on a page; they are a reflection of healthcare access, gender equality, economic stability, and cultural norms.

Through this exploration, we'll gain a deeper understanding of our world and its future, one number at a time.



Plot the US birthrates through time

```{r include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
# Create a function of the above plot code
plot_line <- function(data, iso2c = "US", title = "United States",
                      title_size = 10, hjust = 0,
                      dot_size = 0.6, linewidth = 0.2) {
  # Check if required columns exist in the data
  if (!("iso2c" %in% names(data)) || !("date" %in% names(data)) || !("SP.DYN.TFRT.IN" %in% names(data))) {
    stop("Data must contain 'iso2c', 'date', and 'SP.DYN.TFRT.IN' columns.")
  }

  # Filtering and plotting
  plot <- data %>%
    filter(iso2c == !!iso2c) %>%  # Correct filtering
    ggplot(aes(x = date, y = SP.DYN.TFRT.IN)) +
    geom_line(alpha = 0.7, linewidth = linewidth) +
    geom_point(size = dot_size, alpha = 0.7) +
    geom_smooth(
      span = 0.9,
      linewidth = linewidth,
      linetype = 2,
      se = FALSE,
      color = "#E91E63"
    ) +
    labs(
      x = "",
      y = "",
      title = title
    ) +
    theme(
      plot.title      = element_text(size = title_size, face = "bold",
                                     hjust = hjust),
      axis.title      = element_text(size = 8, face = "bold"),
      axis.text       = element_text(size = 8),
      legend.position = "none",
      panel.grid      = element_blank()
    )

  return(plot)
}

plot_line_small <- function(data, iso2c = "US", title = "United States",
                      title_size = 10, hjust = 0,
                      dot_size = 0.6, linewidth = 0.3) {
  # Check if required columns exist in the data
  if (!("iso2c" %in% names(data)) || !("date" %in% names(data)) || !("SP.DYN.TFRT.IN" %in% names(data))) {
    stop("Data must contain 'iso2c', 'date', and 'SP.DYN.TFRT.IN' columns.")
  }

  # Filtering and plotting
  plot <- data %>%
    filter(iso2c == !!iso2c) %>%  # Correct filtering
    ggplot(aes(x = date, y = SP.DYN.TFRT.IN)) +
    geom_line(alpha = 0.9, linewidth = linewidth) +
    # geom_point(size = dot_size, alpha = 0.7) +
    # geom_smooth(
    #   span = 0.9,
    #   linewidth = linewidth,
    #   linetype = 2,
    #   se = FALSE,
    #   color = "#E91E63"
    # ) +
    labs(
      x = "",
      y = "",
      title = title
    ) +
    theme(
      plot.title      = element_text(size = title_size, face = "bold",
                                     hjust = hjust),
      axis.title      = element_text(size = 8, face = "bold"),
      axis.text       = element_text(size = 7),
      legend.position = "none",
      panel.grid      = element_blank()
    )

  return(plot)
}
```


```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width = "100%", fig.align = "center", fig.height=4}

# United States
line_us <- plot_line_small(fert_tbl, "US", "United States",
                     hjust = 0.5)

# China
line_china <- plot_line_small(fert_tbl, "CN", "China",
                        hjust = 0.5)

# Japan
line_japan <- plot_line_small(fert_tbl, "JP", "Japan",
                        hjust = 0.5)

# Russia
line_russia <- plot_line_small(fert_tbl, "RU", "Russia",
                         hjust = 0.5)

# World
line_world <- plot_line(fert_tbl, "1W", "World",
                        title_size = 14, hjust = 0.5,
                        dot_size = 1.2, linewidth = 0.4)

```

<br>

```{r, echo=FALSE, message=FALSE, warning=FALSE, out.width = "100%", fig.align = "center", fig.height=6}
line_world + 
  (line_us + line_china + line_japan + line_russia) + 
  plot_layout(nrow = 2) +
  plot_annotation(
    caption = "World Bank, World Development Indicators.\n2023. Fertility rate, total (births per woman).\nhttps://data.worldbank.org/indicator/SP.DYN.TFRT.IN",
    theme = theme(plot.caption = element_text(size = 8, color = "gray35"))
  )
```


